---
title: "p8105_hw3_tj2519"
output: github_document
date: "2023-10-09"
---

# Problem 1
## part1: load the library and dataset
```{r}
library(tidyverse)
library(ggridges)
library(ggplot2)
library(p8105.datasets)
data("instacart")

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## part2: calculate the number of aisles and number of items in each aisle
```{r}
problem1_df = instacart 
problem1_df %>% 
  group_by(aisle_id) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs))

view(problem1_df)
```

## part3: make a plot showing the number of items ordered in each aisle
```{r}
problem1_df %>% 
  group_by(aisle_id) %>% 
  summarize(number_item = n()) %>% 
  filter(number_item > 10000) %>% 
  ggplot(aes(x = reorder(aisle_id, -number_item), y = number_item)) +
  geom_bar(stat = "identity", width = 0.5) + 
  theme(axis.text.x = element_text(size = 6)) +
  xlab("aisle_id") 
```

## part4: make a table

```{r}
problem1_df %>% 
  group_by(aisle, product_name) %>% 
  filter(aisle == "baking ingredients" |
           aisle == "dog food care" |
           aisle == "packaged vegetables fruits"
           ) %>% 
  summarize(order_num = n()) %>% 
  arrange(aisle, desc(order_num)) %>% 
  slice(1:3) %>% 
  pivot_wider(names_from = aisle,
              values_from = order_num)

```

## part5: make a new table
```{r}
problem1_df %>% 
  filter(product_name == "Pink Lady Apples" | 
           product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_order_time = mean(order_hour_of_day, na.rm = TRUE)) %>% 
  pivot_wider(names_from = order_dow,
              values_from = mean_order_time)

```



# Problem 2
## part1: load the library and dataset
```{r}
library(p8105.datasets)
data("brfss_smart2010") 

view(brfss_smart2010)
```

## part2: cleaning the data
```{r}
problem2_df = brfss_smart2010

problem2_clean = problem2_df %>% 
  janitor::clean_names() %>% 
  rename(location_abbr = locationabbr, location_desc_order = locationdesc) %>% 
  filter(topic == "Overall Health") %>% 
  drop_na(response) %>% 
  mutate(response = forcats::fct_relevel(
    response, c("Poor", "Fair", "Good", "Very good", "Excellent")))

view(problem2_clean)
```

## part3: answer the first question - the number of locations
```{r}
state_2002 = problem2_clean %>%
  filter(year == 2002) %>% 
  group_by(location_abbr) %>% 
  summarize(n_location = n_distinct(location_desc_order)) %>% 
  filter(n_location >= 7)

problem2_clean %>%
  filter(year == 2002) %>% 
  group_by(location_abbr) %>% 
  summarize(n_location = n_distinct(geo_location)) %>% 
  filter(n_location >= 7)

state_2007 = problem2_clean %>%
  filter(year == 2010) %>% 
  group_by(location_abbr) %>% 
  summarize(n_location = n_distinct(location_desc_order)) %>% 
  filter(n_location >= 7)
```
According to the results, in 2002, `r nrow(state_2002)` states were observed at 7 or more locations (counties), and the 6 states are "`r state_2002$location_abbr`". In 2010, `r nrow(state_2007)` states were observed at 7 or more locations (counties), and the 14 states are "`r state_2007$location_abbr`". 


## part4: answer the second question - the mean data_value for each state
```{r}
problem2_clean %>% 
  filter(response == "Excellent") %>% 
  group_by(location_abbr, year) %>% 
  summarize(data_value_mean = mean(data_value, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = data_value_mean, group = location_abbr, 
             color = location_abbr)) +
  geom_line() +
  theme_minimal() + 
  labs(title = "The change in the mean of data_value for the 'Excellent' category",
       y = "Mean of data value")

```
To make this plot, I move the legend to the right of the plot so it could be more clear. The plot indicates the mean data value (prevalence) change across time (from 2002 to 2010) in each state (with 'Excellent' health). In this plot the mean of data value is calculated and from the plot, it could be observed that the mean fluctuate a lot across time and the trend is also very different among the states.


## part5: answer the third question, make the 2-panel plot showing the data value distribution of the year 2006 and 2010
```{r}
problem2_clean %>% 
  filter(location_abbr == "NY", year == 2006 | year == 2010) %>% 
  ggplot(aes(x = data_value, fill = response)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ year) +
  labs(title = "Data value distribution for the year 2006 and 2010 in New York (density plot)",
       x = "Data value distribution")

problem2_clean %>% 
  filter(location_abbr == "NY", year == 2006 | year == 2010) %>% 
  group_by(year) %>% 
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  facet_grid(. ~ year) +
  labs(title = "Data value distribution for the year 2006 and 2010 in New York (box plot)",
       x = "Data value distribution",
       y = "data value (%)")

```

```{r}
median_2006 = problem2_clean %>% 
  filter(location_abbr == "NY", year == 2006) %>% 
  group_by(response) %>% 
  summarize(median = median(data_value))
  
median_2010 = problem2_clean %>% 
  filter(location_abbr == "NY", year == 2010) %>% 
  group_by(response) %>% 
  summarize(median = median(data_value))
```

I firstly filter the datset to include only the data in NY and only the observations from 2006 and 2010. Then, to show the distribution of each response type, I draw a density plot as well as a box plot. Then, we split the plot according to the value of each year. From both plots, we could tell that in both 2006 and 2010, the data value generally has an increasing trend from 'Poor' to 'Very Good'. However, as for the 'Excellent' group, the data_value decreases. It means that most people in NY self-reported as good health or above in both years. In 2006, the median values for each response type are (from 'Poor' to 'Excellent') `r median_2006$median`, In 2010, the median values for each response type are (from 'Poor' to 'Excellent') `r median_2010$median`.



# Problem 3
## part1: read the two datasets and tidy the two datasets, also apply left_join by "seqn" to merge the two datasets. Also recode certain variables
```{r message = FALSE}
demo_data = 
  read_csv("data/nhanes_covar.csv", skip = 4) %>% 
  janitor::clean_names() %>% 
  filter(age >= 21) %>% 
  drop_na() %>% 
  mutate(sex = recode(sex, `2` = "female", `1` = "male"),
         education = recode(education, `1` = "Less than high school", 
                                      `2` = "High school equivalent",
                                      `3` = "More than high school")) %>% 
  mutate(education = forcats::fct_relevel(
    education, c("Less than high school", "High school equivalent", 
                 "More than high school")))
  

acc_data =
  read_csv("data/nhanes_accel.csv") %>% 
  janitor::clean_names()

demo_acc_join = 
  left_join(demo_data, acc_data, by = "seqn")
```


## part2: make a table indicating the number of each sex in each education level and make a plot showing the distribution of age in the two genders in each education level
```{r}
demo_acc_join %>% 
  group_by(education, sex) %>% 
  summarize(number_count = n()) %>% 
  pivot_wider(names_from = sex, values_from = number_count)

demo_acc_join %>% 
  ggplot(aes(x = sex, y = age)) +
  geom_boxplot() + 
  facet_grid(. ~ education) +
  labs(title = "Age distribution of female and male in each education level")

demo_acc_join %>% 
  ggplot(aes(x = age, fill = sex)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ education)
```
The table indicates that the number of female and male who has a degree of 'Less than high school' is almost equal; among the people with a High school degree or equivalent, the number of female is less than the number of male; among the people with a more than high school degree, the number of female and male is almost equal.

According to the box plot, the median age of female and male with a less than a high school degree is similar, and the age distributions for the two genders are similar; the median age of female with a 'High school equivalent' degree is older than male, and the age distributions for the two genders are similar, with males slightly younger than females; the median age of female with a 'More than high school' degree is similar, and the age distributions for the two genders are similar.


## part3: aggregate the "min" variables to create a single variable, and making a plot
```{r}
demo_acc_join %>% 
  pivot_longer(starts_with("min"), 
               names_to = "activity",
               values_to = "activity_value") %>% 
  group_by(education, sex, age, seqn) %>% 
  summarize(total_activity = sum(activity_value)) %>% 
  ggplot(aes(x = age, y = total_activity, color = sex)) +
  geom_smooth(se = FALSE) +
  geom_point() +
  facet_grid(. ~ education) +
  labs(title = "Total activity across age in each education level")
  
```

The plot shows the total activity across age in each education level, and the trend line in each education level seems to have a trend - with younger age, the total activity is higher. In high school equivalent and more than high school group, female seem to have higher total activity than male. 


## part4: make a plot showing the 24-hr activity for each education level and stratified by sex
```{r message = FALSE}
demo_acc_join %>% 
  pivot_longer(starts_with("min"), 
               names_to = "activity_time",
               values_to = "activity_value") %>%
  mutate(activity_time = substr(activity_time, start = 4, stop = 100)) %>% 
  mutate(activity_time = as.numeric(activity_time)) %>% 
  ggplot(aes(x = activity_time, y = activity_value, color = sex)) + 
  geom_smooth(aes(group = sex), se = FALSE) +
  facet_grid(. ~ education) +
  labs(title = "The activity across 24 hrs in each education level stratified by sex",
       x = "activity time (min)")
```

The plot illustrates the activity across time in each observation (stratified by sex) in each education level.Within the 24-hr period, the activity seems to be the lowest at around the 250th min in each education group, and the highest activity seem to appear at around the 750th min in each education group. Female and male shows similar trend in each education level, with 'Less than high school' group showing higher activity around the 750th min.


